A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-07-10 text/manuscript.Rmd
added more code for fig4 - will be complete tomorrow, table 1, and table 2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(foreach)
library(sqldf)
library(tidyverse)
library(ggthemes)
library(data.table)
library(biomaRt)
library(dplyr)
library(GenomicRanges)
library(SummarizedExperiment)
library(reactome.db)
library(plyr)
library(cluster)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(umap)
library(dbscan)
library(class)
library(TCGAbiolinks)
library(patchwork)
library(ggplotify)
library(ComplexHeatmap)
library(knitr)
library(kableExtra)
library(UpSetR)
```

@@

save(path_match_in_set, file = "path_match_in_set.Rda")

@@

save(filtered_mut_data, gene_order, file = "filtered_mut_data.Rda")
pheatmap(filtered_mut_data[gene_order,],
         show_rownames = T,
         show_colnames = F,
         treeheight_row = 0,
         treeheight_col = 0,
         cellheight = 5,
         cellwidth = 1,
         cluster_cols = T,
         cluster_rows = F,
         #gaps_row = file_dims,
         #cutree_cols = 6,
         annotation_legend = F,
         legend_breaks = c(0, 0.25, 0.75, 1),
         legend_labels = c("","Not Mut", "Mut", ""),
         color = c("grey", "firebrick3"),
         clustering_method = "ward.D2",
         fontsize = 10,
         legend = T,
         filename = "fig4.png")

## stopped here ## ignore bottom using as reference
path_match_in_set
cancer_heatmap <- pheatmap(cancer_path[pathway_dendro,], ## using pathway row order 
                          show_rownames = F,
                          show_colnames = F,
                          cellheight = 1,
                          cellwidth = 0.5,
                          cluster_cols = T,
                          cluster_rows = F,
                          annotation_legend = F,
                          color = c("grey", "firebrick3"),
                          clustering_method = "single", ## ordering is single linkage
                          legend = T)
  cancer_order <- cancer_heatmap$tree_col$order ## cancer order
  cancer_path <- names(cancer_path) ## getting the sample_id original order 
  cancer_dendro[[n]] <- cancer_path[cancer_order] ## reorder cancer by heatmap order
  rm(cancer_info, cancer_path)
  gc()
}
dev.off()

cancer_disrupt_heatmap <- left_join(cancer_disrupt_heatmap, cancer_order, by = "sample_id") ## add in cancer type information for faceting
save(cancer_disrupt_heatmap, file = "cancer_disrupt_heatmap.Rda") 
```

```{r figure-4}

@@

<mark>As discussed previously, prior efforts to extract common signatures from
pan-cancer datasets met with difficulty in distinguishing tumor samples from
their tissue-specific -omics data signatures. Given our pathway-disruption based
classification, this raises the question, are tumor phenotypes driven entirely
by common driver genes, or by "silent" tissue-specific effectors ( _i.e._ too
few samples to detect above statistical significance thresholds), or a
combination of both? To answer this question, we ranked the top cluster-specific
pathways for each cluster, and then tabulated the most frequently mutated genes
for each pathway as a function of cancer type (**Table 1** and **Supplementary
tables ST2-ST13**). We find that in most clusters the top pathways are
determined by mutations in a handful of driver genes, TP53, PIK3CA, KRAS etc.
However in many cancers we find divergence outside these top driver mutations,
wherein etc etc</mark>

![](../figures/table1.png)
<p style="FONT-SIZE:10px; COLOR:#000000; LINE-HEIGHT:12px;">**Table 1:** The numbers in this table are completely wrong, it is just a place holder, please ignore</p>

@@

Therefore, we calculated cluster-specific enrichment in metastatic tumors and
found 99 enriched pathways across all clusters (**table 2**). A number of
enrichments were found in multiple clusters represented by pathways that were
already shown to be enriched in a neighboring cluster. For example,
“Beta-catenin independent WNT signaling” is enriched in non-metastatic samples
of clusters 6, 7, 11 and 12 (see **supplementary figure S4**) but not in 9 and
10. This pathway is enriched in metastatic tumors of clusters 9 and 10, plus the
union of 9+10 (p < 10-3, **table 2**). This is also true of “Erythropoietin
activates RAS”, which is enriched in clusters 11 and 12 (**supplementary figure
S4**) and also in metastatic tumors of clusters 9 and 10, plus cluster 7.
Cluster 1, one of two clusters for which we could not identify any
distinguishing pathways (cluster 8 being the other), had enrichment in two
pathways that were otherwise specific to non-metastatic tumors of other
clusters. These include “Cellular response to hypoxia” (enriched in clusters 6 &
7) and “TNFR2 non-canonical NF-kB pathway” (cluster 6 & 7). Cluster 5 metastatic
samples were enriched for two key neutrophil pathways, “Neutrophil
Degranulation” and “Fc epsilon receptor (FCERI) signaling” which are specific to
cluster 9 & 10 non-metastatic tumors. Thus, pathway lesions specific to
metastases tend toward complementary mutations that are otherwise specific to
non-metastatic tumors of other clusters.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(rstan)
library(foreach)
library(sqldf)
library(tidyverse)
library(ggthemes)
library(data.table)
library(biomaRt)
library(dplyr)
library(GenomicRanges)
library(SummarizedExperiment)
library(reactome.db)
library(plyr)
library(cluster)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(umap)
library(dbscan)
library(class)
library(TCGAbiolinks)
library(patchwork)
library(ggplotify)
library(ComplexHeatmap)
library(knitr)
library(kableExtra)
library(UpSetR)
```

@@

save(path_match_in_set, file = "path_match_in_set.Rda")
load("../data/path_match_in_set.Rda")

@@

save(filtered_mut_data, gene_order, file = "filtered_mut_data.Rda")

fig4_pheatmap <- pheatmap(filtered_mut_data[gene_order,],
         show_rownames = T,
         show_colnames = F,
         treeheight_row = 0,
         treeheight_col = 0,
         cellheight = 5,
         cellwidth = 1,
         cluster_cols = T,
         cluster_rows = F, ## turn off row clustering
         annotation_legend = F,
         legend_breaks = c(0, 0.25, 0.75, 1),
         legend_labels = c("","Not Mut", "Mut", ""),
         color = c("grey", "firebrick3"),
         clustering_method = "ward.D2",
         fontsize = 10,
         legend = T)

## getting sample orders
pheatmap_order <- fig4_pheatmap$tree_col$order ## getting sample clustering order
sample_order <- names(filtered_mut_data) ## getting the sample_id original order 
sample_order <- sample_order[pheatmap_order] ## reorder cancer by heatmap order

## stopped here ## ignore bottom using as reference

class67_gene_2_path_heatmap <- filtered_mut_data %>% rownames_to_column(var = "gene_name") %>%  ## move rownames to column
  gather(., key = "sample_id", value = "membership", -gene_name) ## wide to long format
class67_gene_2_path_heatmap$membership <- as.character(class67_gene_2_path_heatmap$membership) ## turn boolean values into categories

## set levels based on clustering algorithm above
class67_gene_2_path_heatmap$gene_name <- factor(class67_gene_2_path_heatmap$gene_name, levels = c(rev(rownames(filtered_mut_data))))
class67_gene_2_path_heatmap$sample_id <- factor(class67_gene_2_path_heatmap$sample_id, levels = c(sample_order))

class67_gene_2_path_heatmap <- left_join(class67_gene_2_path_heatmap, path_match_in_set, by = "gene_name") ## add in path_combo for faceting 
save(class67_gene_2_path_heatmap, file = "class67_gene_2_path_heatmap.Rda") 
rm(list = ls(all.names = T))
gc()
```

```{r figure-4, eval = FALSE}
load("../data/class67_gene_2_path_heatmap.Rda") ## load data
load("../data/path_match_in_set.Rda")

### getting facet ordering
most_genes <- path_match_in_set %>% .$path_combo %>% table() %>% as.data.frame() %>% arrange(desc(Freq)) %>% ## arranging pathway combination based on most genes involved
  'colnames<-' (c("path_combo", "Freq")) %>% 
  rownames_to_column(var = "order")

fig4 <- ggplot(data = class67_gene_2_path_heatmap, aes(y=gene_name, x=sample_id, fill = membership)) + 
  geom_tile(size = 0.6) +
  scale_fill_manual(values = c("grey", "firebrick3")) +
  #facet_grid(factor(path_combo, levels = most_genes$path_combo) ~ ., scales = "free", space = "free") +
  #scale_y_discrete(position = "left") +
  labs(tag = "A") +
  ylab("") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(),
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        legend.position = "none") 
fig4
ggsave("fig4_ggplot_heatmap.png", height = 100, width = 17, limitsize = FALSE)

@@

<mark>As discussed previously, prior efforts to extract common signatures from
pan-cancer datasets met with difficulty in distinguishing tumor samples from
their tissue-specific -omics data signatures. Given our pathway-disruption based
classification, this raises the question, are tumor phenotypes driven entirely
by common driver genes, or by "silent" tissue-specific effectors ( _i.e._ too
few samples to detect above statistical significance thresholds), or a
combination of both? To answer this question, we ranked the top cluster-specific
pathways for each cluster, and then tabulated the most frequently mutated genes
for each pathway as a function of cancer type (**Table 1** and **Supplementary
tables ST2-ST13**). We find that in most clusters the top pathways are
determined by mutations in a handful of driver genes, TP53, PIK3CA, KRAS etc.
However in many cancers we find divergence outside these top driver mutations,
wherein etc etc</mark>
<div style="text-align: right"> Table 1 Data </div>
```{r table-1-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_2d_3d_coors.tsv") ## load in data
load("../data/moderate_enrich_prob.Rda") ## load data
load("../data/gene2_level2_lookup.Rda")
load("../data/gene_data.Rda")
load("../data/gdc_gene_exp_adj.Rda")
load("../data/pathname2level2.Rda")

## parameters ##
enrich_cutoff <- 0.30
howmanypaths <- 5 ## top _ pathways; how many top pathways do you want to see
show_perc <- TRUE
show_n <- TRUE
mutated_gene_position <- 1
################

clust_spec_enrich <- moderate_enrich_prob %>% filter(signif <= 2500 & difference >= enrich_cutoff) ## signif is 95% CI with a specified cutoff

clust_spec_enrich_top5 <- list()
for(n in unique(clust_spec_enrich$clust)) {
  message(n)
  clust_spec_enrich_top5[[n]] <- clust_spec_enrich %>% filter(clust == n) %>% ## filter for specific cluster
    arrange(desc(difference)) %>% ## arrange enrichment score in decreasing order 
    dplyr::select(path, clust) %>% .[1:5,]
}
clust_spec_enrich_top5 <- clust_spec_enrich_top5 %>% do.call(rbind, .) %>% as.data.frame() ## combine list from for loop
rm(clust_spec_enrich)

## finding the top mutated gene for each cluster in each pathway for the top 5 pathways 
type <- unique(tumor_clusters$project_code) ## type of cancer type in data
cluster_spec_enrich_table <- list() ## list for forloop

for(d in unique(clust_spec_enrich_top5$clust)) {
  message(d)
  clust_toppath <- clust_spec_enrich_top5 %>% filter(clust == d) %>% .$path %>% as.character() ## get the top pathway names for spec class
  class_spec <- tumor_clusters %>% dplyr::select(V1, knn_clust, project_code) %>% filter(knn_clust %in% d) ## get the smaple ids and cancer type for spec class
  top_genes_in_path_cancer_type <- list()
  for(n in clust_toppath) { ## for each of the top paths do this
    message(n)
    gene_in_path <- id_to_names %>% .[,which(names(.) == n)] %>% as.data.frame() %>% ## select pathway
    'rownames<-' (rownames(id_to_names)) %>% 
    filter(. == 1) %>% ## filter for genes that are in that pathway
    rownames(.)
    top_gene_in_path <- list()
    for(t in type) { ## looping through cancer type within class
      message(t)
      class_spec <- tumor_clusters %>% dplyr::select(V1, knn_clust, project_code) %>% filter(knn_clust == d) ## sample_id for class
      if(t %in% class_spec$project_code) { ## if cancer type is in class then run this 
        class_type_spec <- class_spec %>% filter(project_code == t) %>% column_to_rownames(var = "V1") %>% rownames() ## get cancer type specific tumor classes
      if(length(class_type_spec) == 1) { ## if you only have 1 sample_id in the class
        top_mut_gene <- gene_mutation %>% rownames_to_column(var = "gene") %>% .[.$gene %in% gene_in_path, names(.) %in% c("gene",  class_type_spec)] %>% remove_rownames() %>% column_to_rownames(var = "gene") %>% 'colnames<-' ("obs_count") %>% 
          mutate(n_sample = length(class_type_spec),
               perc_mut = round((obs_count / length(class_type_spec)) * 100, digits = 1)) %>% 
        rownames_to_column(var = "gene") %>% arrange(desc(obs_count))
      }
      if(length(class_type_spec) > 1) { ## if you have more than 1 sample_id in the class
        top_mut_gene <- gene_mutation %>% .[rownames(.) %in% gene_in_path, names(.) %in% class_type_spec] %>% 
        rowSums() %>% as.data.frame() %>% 'colnames<-' ("obs_count") %>%
        mutate(n_sample = length(class_type_spec),
               perc_mut = round((obs_count / length(class_type_spec)) * 100, digits = 1)) %>% 
        rownames_to_column(var = "gene") %>% arrange(desc(obs_count))
      }
      if(show_perc == TRUE & show_n == FALSE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position], " (", top_mut_gene$perc_mut[mutated_gene_position],"%)")  
      }
      if(show_perc == TRUE & show_n == TRUE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position], " (", top_mut_gene$perc_mut[mutated_gene_position],"%)", "    n = ", length(class_type_spec))  
      }
      if(show_perc == FALSE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position])  
      }
      }
      if(!(t %in% class_spec$project_code)){ ## if cancer type is not in class then insert NA
        top_gene_in_path[[t]] <- "NA" 
      }
    }
    top_gene_in_path <- top_gene_in_path %>% do.call(rbind, .) %>% as.data.frame() %>% 'colnames<-' (n)
    clust_row <- data.frame(Class = d, row.names = n) %>% t() ## creating a row for cluster membership
    top_genes_in_path_cancer_type[[n]] <- rbind(clust_row, top_gene_in_path)
  }
  cluster_spec_enrich_table[[n]] <- top_genes_in_path_cancer_type %>% do.call(cbind, .) %>% as.data.frame() %>% 
  t() %>% as.data.frame() %>% rownames_to_column(var = "Pathway")
  }

cluster_spec_enrich_table1 <- cluster_spec_enrich_table %>% do.call(rbind, .) %>% as.data.frame()
save(cluster_spec_enrich_table1, file = "cluster_spec_enrich_table1.Rda")
rm(list = ls(all.names = T))
gc()
```

```{r table-1}
load("../data/cluster_spec_enrich_table1.Rda")
cluster_spec_enrich_table1 %>% kable(caption = "Table 1: Most mutated gene in class-specific pathways (enrichment score >= 0.30) for specific cancer", row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = T)
```
<p style="FONT-SIZE:10px; COLOR:#000000; LINE-HEIGHT:12px;">**Table 1:** The numbers in this table are completely wrong, it is just a place holder, please ignore</p>

@@

Therefore, we calculated cluster-specific enrichment in metastatic tumors and
found 99 enriched pathways across all clusters (**table 2**). A number of
enrichments were found in multiple clusters represented by pathways that were
already shown to be enriched in a neighboring cluster. For example,
“Beta-catenin independent WNT signaling” is enriched in non-metastatic samples
of clusters 6, 7, 11 and 12 (see **supplementary figure S4**) but not in 9 and
10. This pathway is enriched in metastatic tumors of clusters 9 and 10, plus the
union of 9+10 (p < 10-3, **table 2**). This is also true of “Erythropoietin
activates RAS”, which is enriched in clusters 11 and 12 (**supplementary figure
S4**) and also in metastatic tumors of clusters 9 and 10, plus cluster 7.
Cluster 1, one of two clusters for which we could not identify any
distinguishing pathways (cluster 8 being the other), had enrichment in two
pathways that were otherwise specific to non-metastatic tumors of other
clusters. These include “Cellular response to hypoxia” (enriched in clusters 6 &
7) and “TNFR2 non-canonical NF-kB pathway” (cluster 6 & 7). Cluster 5 metastatic
samples were enriched for two key neutrophil pathways, “Neutrophil
Degranulation” and “Fc epsilon receptor (FCERI) signaling” which are specific to
cluster 9 & 10 non-metastatic tumors. Thus, pathway lesions specific to
metastases tend toward complementary mutations that are otherwise specific to
non-metastatic tumors of other clusters.
<div style="text-align: right"> Table 2 Data </div>
```{r table-2-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_2d_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

tumor_clusters <- tumor_clusters %>% dplyr::select(V1, ajcc_metastasis_pathologic_pm, knn_clust)

m0_clust <- tumor_clusters[tumor_clusters$ajcc_metastasis_pathologic_pm == "M0",] ## filter for nonmetastatic samples
m1_clust <- tumor_clusters[tumor_clusters$ajcc_metastasis_pathologic_pm == "M1",] ## filter for metastatic samples

prepare_rbeta_data_p1 <- function(data, path_analysis) {
  n_tumor_in_group <- as.data.frame(table(data$ajcc_metastasis_pathologic_pm)) ## Getting the total number of tumor samples in each group
  path_analysis <- path_analysis[,names(path_analysis) %in% data$V1] ## Filter only group specific cancers
  data <- data[match(names(path_analysis), data$V1),] ## Match tumor sample ids
  table(names(path_analysis) == data$V1) ## good they match now
  ## Now we are going to build group specific cluster data
  n_tumor_in_clust <- as.data.frame(table(data$knn_clust)) ## getting total n in each cluster
  names(n_tumor_in_clust) <- c("clust", "tot_n_clust") ## rename colnames
  clust_group <- as.character(n_tumor_in_clust$clust) ## Getting the cluster numbers
  cluster_path_count <- list() ## Creating a list for the forloop
  for(n in clust_group) {
    print(n)
    df <- subset(data, knn_clust == n) ## subset cluster number
    df <- df$V1 ## getting tumor sample id
    df1 <- path_analysis[,names(path_analysis) %in% df] ## getting the tumor samples from path analysis
    df1 <- as.data.frame(rowSums(df1))
    names(df1) <- n
    cluster_path_count[[n]] <- df1
  }
  cluster_path_count <- do.call(cbind, cluster_path_count) ## Combine the list together
  rbeta_path_df <- cluster_path_count %>% rownames_to_column(var="path") 
  rbeta_path_df <- rbeta_path_df %>% gather(key = "clust", value = "mutation", -path) ## Wide to long format
  global_path_mut <- as.data.frame(rowSums(cluster_path_count))
  global_path_mut <- global_path_mut[rep(1:ncol(global_path_mut), each = max(as.numeric(as.character(rbeta_path_df$clust))))] ## Repeat column 12 times
  names(global_path_mut) <- as.character(clust_group) ## Replace names with cluster numbers
  global_path_mut <- global_path_mut %>% rownames_to_column(var="path") %>% gather( key = "clust", value = "global", -path) ## Wide to long format
  rbeta_path_df <- left_join(rbeta_path_df, global_path_mut, by = c("path", "clust")) ## left_join rbetadf and global
  rbeta_path_df <- left_join(rbeta_path_df, n_tumor_in_clust, by = "clust")
  rbeta_path_df$global_sample <- as.numeric(sum(n_tumor_in_clust$tot_n_clust)) ## Getting total tumor sample number
  names(rbeta_path_df) <- c("path", "clust", "clust_mut", "global_mut", "tot_n_clust", "global_sample")
  rbeta_path_df$tot_n_clust <- as.numeric(as.character(rbeta_path_df$tot_n_clust)) ## convert tot_n_clust numeric
  return(rbeta_path_df)
}

m0_rbeta <- prepare_rbeta_data_p1(data = m0_clust, path_analysis = gdc_reactome_path_analysis)
## Combining cluster 6/7 cluster 910 cluster 1112
combo_group <- c("6", "9", "11")
compo_group_df <- list() ## list for loop
for(n in combo_group) {
  print(n) ## first cluster in set
  m <- as.numeric(n) + 1 ## second cluster in set
  m <- as.character(m)
  df <- m0_rbeta[m0_rbeta$clust == n,] ## get rbeta info for first cluster in set
  df1 <- m0_rbeta[m0_rbeta$clust == m,] ## get rbeta info for second cluster in set
  tot_clust_mut <- df$clust_mut + df1$clust_mut ## add both clust muts
  tot_n_clust <- df$tot_n_clust + df1$tot_n_clust ## add both clust n's
  p <- paste0(n,m)
  df2 <- m0_rbeta[1:377,]
  df2$clust <- p ## replace with new clust name
  df2$clust_mut <- tot_clust_mut ## replace with new clust mut count
  df2$tot_n_clust <- tot_n_clust ## replace with new clust size
  compo_group_df[[p]] <- df2 ## store in list and continue loop
}
compo_group_df <- do.call(rbind, compo_group_df) ## rbind list
m0_rbeta <- rbind(m0_rbeta, compo_group_df) ## rbind m0_rbeta and the compo group together
save(m0_rbeta, file = "m0_rbeta.Rda")

## repeat same function above with metastasis sample
m1_rbeta <- prepare_rbeta_data_p1(data = m1_clust, path_analysis = gdc_reactome_path_analysis)
compo_group_df <- list() ## list for loop
for(n in combo_group) {
  print(n) ## first cluster in set
  m <- as.numeric(n) + 1 ## second cluster in set
  m <- as.character(m)
  df <- m0_rbeta[m0_rbeta$clust == n,] ## get rbeta info for first cluster in set
  df1 <- m0_rbeta[m0_rbeta$clust == m,] ## get rbeta info for second cluster in set
  tot_clust_mut <- df$clust_mut + df1$clust_mut ## add both clust muts
  tot_n_clust <- df$tot_n_clust + df1$tot_n_clust ## add both clust n's
  p <- paste0(n,m)
  df2 <- m0_rbeta[1:377,]
  df2$clust <- p ## replace with new clust name
  df2$clust_mut <- tot_clust_mut ## replace with new clust mut count
  df2$tot_n_clust <- tot_n_clust ## replace with new clust size
  compo_group_df[[p]] <- df2 ## store in list and continue loop
}
compo_group_df <- do.call(rbind, compo_group_df) ## rbind list
m1_rbeta <- rbind(m1_rbeta, compo_group_df) ## Combine the new groups
save(m1_rbeta, file = "m1_rbeta.Rda")

## function to compiling data for rbeta analysis
prepare_rbeta_data_p2 <- function(data) {
  enrich_prob <- data.frame(path = data$path,
                               clust = data$clust,
                               fg_mut = data$clust_mut,
                               fg_not_mut = data$tot_n_clust - data$clust_mut,
                               bg_mut = data$global_mut,
                               bg_not_mut = data$global_sample - data$global_mut)
fg_mut_rate <- list()
bg_mut_rate <- list()
for(n in unique(as.character(enrich_prob$clust))) {
  print(n)
  df <- subset(enrich_prob, clust == n)
  fg_denominator <- nrow(df) * (df$fg_mut[1] + df$fg_not_mut[1]) ## n_path * tot_n_clust_x
  fg_numerator <- sum(df$fg_mut)
  bg_denominator <- nrow(df) * (df$bg_mut[1] + df$bg_not_mut[1])
  bg_numerator <- sum(df$bg_mut)
  fg_mut_rate[[n]] <- fg_numerator / fg_denominator
  bg_mut_rate[[n]] <- bg_numerator / bg_denominator
}

fg_mut_rate <- as.data.frame(do.call(rbind, fg_mut_rate)) ## rbind the mut rates from list
bg_mut_rate <- as.data.frame(do.call(rbind, bg_mut_rate)) ## rbind the mut rates from list
enrich_prob_df <- enrich_prob
enrich_prob_df$fg_mut_rate <- rep(x = fg_mut_rate$V1, each = length(unique(enrich_prob_df$path)))
enrich_prob_df$bg_mut_rate <- rep(x = bg_mut_rate$V1, each = length(unique(enrich_prob_df$path)))
enrich_prob_df$rel_odds <- enrich_prob_df$fg_mut_rate / enrich_prob_df$bg_mut_rate
enrich_prob_df$denominator <- (enrich_prob_df$bg_mut * enrich_prob_df$rel_odds) +   (enrich_prob_df$bg_not_mut / enrich_prob_df$rel_odds)
enrich_prob_df$x_adj <- round((enrich_prob_df$fg_mut + enrich_prob_df$fg_not_mut) * (enrich_prob_df$bg_mut * enrich_prob_df$rel_odds) / enrich_prob_df$denominator)
enrich_prob_df$y_adj <- (enrich_prob_df$fg_mut + enrich_prob_df$fg_not_mut) -   enrich_prob_df$x_adj
return(enrich_prob_df)
}

m0_rbeta <- prepare_rbeta_data_p2(m0_rbeta)                               
m1_rbeta <- prepare_rbeta_data_p2(m1_rbeta)

save(m0_rbeta, m1_rbeta, file = "m0_m1_rbeta.Rda")
rm(list = ls(all.names = T)) ## cleaning up data
gc()
load("../data/m0_m1_rbeta.Rda")
## Metastatic (foreground) - Nonmetastatic (background) Rbeta
moderate_fg <- apply(m1_rbeta[,c("fg_mut", "fg_not_mut")], MARGIN = 1, function(x) {
  rbeta(n = 1e5, shape1=x["fg_mut"], shape2=x["fg_not_mut"])
})
moderate_bg <- apply(m0_rbeta[,c("fg_mut", "fg_not_mut")], MARGIN = 1, function(x) {
  rbeta(n = 1e5, shape1=x["fg_mut"], shape2=x["fg_not_mut"])
})

metastasis_cluster_moderate <- moderate_fg - moderate_bg  ## Fg - Bg ... to see how many times the Fg is larger than the Bg
metastasis_cluster_moderate <- as.data.frame(metastasis_cluster_moderate)

## ModerateEnrichment probability 
prob <- moderate_fg > moderate_bg ## boolean data
mean_prob <- apply(prob, MARGIN = 2, function(x) {
  mean(x)
})
## Moderate Difference
CI <- 0.025

diff_fg_bg <- moderate_fg - moderate_bg ## fg - bg
rm(moderate_fg, moderate_bg, prob)
gc()
quant_diff <- t(apply(diff_fg_bg, MARGIN = 2, function(x) {
  quantile(x, c(CI, 0.5, 1-CI)) ## 95% CI
}))

sig_fg_bg <- as.data.frame(apply(diff_fg_bg, MARGIN = 2, function(x) {
  sum(x < 0 ) ## out of 1e5 simulations, how many times is fg-bg less than 0
}))

m1_minus_m0_per_cluster_rbeta <- data.frame(path = m1_rbeta$path,
                                   clust = m1_rbeta$clust,
                                   probability = mean_prob,
                                   signif = sig_fg_bg[,1],
                                   difference = quant_diff[,2],
                                   lower = quant_diff[,1],
                                   upper = quant_diff[,3])
save(m1_minus_m0_per_cluster_rbeta, file = "m1_minus_m0_per_cluster_rbeta.Rda")
rm(list = ls(all.names = T))
gc()
```

```{r table-2}
load("../data/m1_minus_m0_per_cluster_rbeta.Rda")

table2 <- m1_minus_m0_per_cluster_rbeta %>% filter(difference >= 0.20 & signif <= 2500) %>% ## filter effect size >= 0.20 and enrichment at a 95% CI
  dplyr::select(path, clust, signif, difference) %>% 'colnames<-' (c("Pathway", "Cluster", "Signif", "Enrichment_Score")) ## select relevant column and rename col

table2$log10pvalue <- round(-log10((table2$Signif + 1) / 1E5), digits = 2) ## Getting p value ## adding 1 for pseudo count.. 1E5 is the n simulations performed
table2$Enrichment_Score <- round(table2$Enrichment_Score, digits = 2)
table2$Signif <- NULL ## we don't need this anymore

table2 %>% knitr::kable(format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = T, color = "black")
```
